---
title: "TimeSeries Assignment11"
author: "Venkat Kasarla"
date: "11/30/2017"  
output: 
    html_document:
          keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dygraphs")
library("forecast")
library("fpp2")
library("xts")
library("dplyr")
```
## Presented by  Venkat Kasarla 

<br>

# Introduction
Present the answrs to the TimeSeries data questions for time series data sets.

<br>

###  1.a Pull the DAX  index

```{r missmatch, echo=TRUE}
help(EuStockMarkets) 
dax <- EuStockMarkets[,"DAX"]
```

<br>

### 1.b plot data and   add a vertical red line to your plot which divides pre-1997 and post-1997 


```{r daxIndex}
plot(dax,main="Stock Data", sub="DAX Index",xlab="Year", ylab="Stock Value")
abline(v=1997, col="red")
plot(EuStockMarkets,main="European Stock Data",xlab="Year", ylab="Stock Value")
abline(v=1997, col="red")
```


<br>

### 1.c Decompose the time series into its components


```{r Decompose}
dax.comp <- decompose(dax, type="mult")
plot(dax.comp)
#HoltWinters filtering on indices
dax.hw <- HoltWinters(dax)

plot(dax.hw,main="Stock Data", sub="DAX Index",xlab="Year", ylab="Stock Value")
```


<br>

### 2.a  load fpp2 and look at ffp2 temperature dataset


```{r  loadfpp2 }
help(maxtemp)
??maxtemp
```

### 2.b Subset after 1990 year


```{r subsetfrom1991}
temp.data<-window(maxtemp, start=1991,initial=c("optimal"))
```

<br>

### c SES to predict the next five years of maximum temperatures in Melbourne. plot this infromation


```{r SES}
plot(temp.data,main="Max Temp Data",xlab="Year", ylab="max tempValue")
fcast.ses <- ses(temp.data,h=5)
plot(fcast.ses,main="Max Temp Data",xlab="Year", ylab="max tempValue")
```

<br>

### 2.d damped Holt’s linear trend to also predict out five years


```{r holt}
fcast.holt <- holt(temp.data,h=5,initial=c("optimal"),damped=TRUE)
plot(fcast.holt,main="Max Temp Data",xlab="Year", ylab="max tempValue")
```


<br>


### 2.e Compare the AICc - Best value in models is to choose the one with the minimum AIC value.SES perefed as it has Low AICc value


```{r AICc}
summary(fcast.ses)
summary(fcast.holt)
```

### 3.a Read the files and rename columns to meaningful name


```{r readRename}
ts.gregorovitch <-  read.csv('../data/Unit11TimeSeries_Gregorovitch.csv')
ts.ollivander <-  read.csv('../data/Unit11TimeSeries_Ollivander.csv')

colnames(ts.gregorovitch) <- c("Date","WandsSold")
colnames(ts.ollivander) <- c("Date","WandsSold")
```

<br>

### 3.b covert to Date class


```{r converDate}
ts.gregorovitch$Date <- as.Date(ts.gregorovitch$Date , "%m/%d/%y")
ts.ollivander$Date <- as.Date(ts.ollivander$Date , "%m/%d/%y")
```

<br>

### 3.c make each data frame an xts object 


```{r createXts}
greg.xts <- xts(ts.gregorovitch$WandsSold,ts.gregorovitch$Date,order.by = ts.gregorovitch$Date )
olliv.xts <- xts(ts.ollivander$WandsSold,ts.ollivander$Date,order.by = ts.ollivander$Date )
```

<br>

### 3.d Bind the two xts objects together and create a dygraph


```{r dygraph}
ollivander <- ts(log(olliv.xts), start = start(olliv.xts), end = end(olliv.xts), frequency = 1)
gregorovitch <- ts(log(greg.xts), start = start(greg.xts), end = end(greg.xts), frequency = 1)
bind.dyg <- cbind(ollivander,gregorovitch)

dygraph(bind.dyg) %>% dyRangeSelector(height=100) 

dygraph(ollivander) %>% dyRangeSelector(height=100) %>% dyShading(from = "1955-1-1", to = "1959-1-1", color = "#CCEBD6")
```

<br>


<br><br><br>
